\chapter*{Acknowledgements}
This work was funded by the Wellcome Trust (Grant Reference: 215153/Z/18/Z). For the purpose of open access, I have applied a CC BY public copyright licence to any Author Accepted Manuscript version arising from this document. I would like to thank the Wellcome Trust for the financial support, networking support, professional development opportunities, and general support given to me during the course of my PhD programme.

I would like to extend thanks to my academic support network. To Stephen Eglen for his guidance, support, and scientific insights throughout the entirety of my PhD programme. To Kristian Franze for his scientific insights, discussion, and opportunities to engage with the Franze lab as well those in the lab. To David Willshaw for being an engaging, insightful, and understanding collaborator as well as hosting me in the Informatics School at Edinburgh. To the Mathematical Genomics and Medicine directors, staff, and students for offering support in a multitude of ways in addition to providing many opportunities for development. To Peter Robinson for hosting me in his lab and the scientific discussion that followed. To the Department of Applied Mathematics and Theoretical Physics for their administrative support throughout the project. To Trinity College for the lively academic and social environment, accommodation, and pastoral support they have offered. 

I would like to also extend thanks to my friends and family. Vira Koshkina for being with me at every step of the way and offering support, reassurance, encouragement, and love --- thank you. To my parents, brother, and family who have been unconditionally supportive of my academic career at all stages and I would not have made it here without them. To all the friends I made in Cambridge (The Byron's Bear Cubs, those in the MGM, those in College and the department, and those met along the way), and to my friends now in far-flung places all around the world, thank you for the cheer, motivation, and all-round good times.
 
